Notes on area-length problem for CLE

### 2014-03-02 area-length/notes.tex
Minor changes
Uniformization theorem added to the notes, bibliography update, paper started
A lot of changes in this commit. I've added text on uniformization theorem, though it is not complete yet. I need to write about the connection with Liouville equation. Some mock up of a paper is also added, but it does not contain new material, just some text copied from notes.

Some results exist in literature for particular case of self-avoiding walks ($\kappa=\frac{8}{3}$). We start
with the description of these results \ref{sec:self-avoid-polyg}, then show that general problem
leads to Liouville field theory \ref{sec:scaling-random-loops} and obtain our result using Liouville
field theory fusion rules \ref{sec:liouville-theory}.

@@

\subsection{$O(n)$ model}
\label{sec:introduction-1}

Self-avoiding polygons can be thought as the limit $n\to 0$ of loops in $O(n)$ model.

Partition function for $O(n)$ model is
\begin{equation}
  \label{eq:9}
  Z_{O(n)} = \sum_{\mbox{loop conf}} x^{\mbox{total length}}
  n^{\mbox{number of loops}}
\end{equation}
Critical point is $x=\frac{1}{\sqrt{2+\sqrt{2-n}}}$, the continuum
limit in this point is described by minimal model $(p,p')$ with
$n=-2\cos\left(\frac{\pi p}{p'}\right)$. $c=1-6\frac{(p-p')^2}{pp'}$, $b=\sqrt{\frac{3}{2}}$.

Self-avoiding walk is obtained in the limit $n\to 0$. Then $c=0$ and
we have logarithmic conformal field theory.

Logarithmic CFT $\sim$ Liouville theory ???

\section{Scaling of random loops}
\label{sec:scaling-random-loops}

Cardy \cite{cardy2001exact} argues that generating function $G^{(r)}$ for rooted self-avoiding loops is given by a derivative w.r.t. $x$
of expectation value of $e^{-gA}$ (eq.6):
\begin{equation}
  \label{eq:82}
  Z=\left<e^{-gA}\right>_{\mathrm{SAL}}=\left<e^{-\sqrt{g}\int J_{\lambda}A_{\lambda}d^{2}x}\right>_{\mathrm{SAL}.A}
\end{equation}
But reasoning is quite strange and unsound. Due to the need to use $g$ for metric in later sections
we will denote area coupling constant by $\mu$ and loop length coupling constant by $\mu_{B}=-\ln x$
(boundary).

In this section we introduce scaling function for random loops.

Consider rooted loops in $O(n)$ model. Let $N^r(L)$ be number of
rooted loops with a given length.  By $N^r(L,A)$ we denote the number
of loops of given length and area. $N^r(L)=\int_0^{\infty} N^r(L,A)
dA$. Area weighted number of loops is denoted by $G^r(L,\mu)$. (Here
$\mu$ is bulk coupling constant ).
\begin{equation}
  \label{eq:2}
  G^r(L,\mu)=\int e^{\mu A} N^r(L,A) dA
\end{equation}

The idea is to write $G$ as a correlation function in Liouville
theory.

@@

@@

Partition function is product over loops of different length
\begin{equation}
  \label{eq:10}
  Z_{O(n)} = \sum_{\mbox{loop conf}} \prod_L x^{L N(L)} n^{N(L)} = \sum_{\mbox{loop conf}} \prod_L\prod_A x^{L N(L,A)} n^{N(L,A)}
\end{equation}

We can rewrite $G^r(\mu,x)$ as the derivative of generating function
\begin{equation}
  \label{eq:7}
  G^r(\mu,x)=\mu\partial_x \log Z(\mu,\mu_B)
\end{equation}

Here
\begin{equation}
  \label{eq:8}
  Z(\mu,\mu_B) = \left< e^{-\mu A} \right> = \frac{\sum_{\mbox{loop conf}}
  \prod_L\prod_A x^{L N(L,A)} n^{N(L,A)} e^{-\mu A}}{\sum_{\mbox{loop conf}} \prod_L\prod_A x^{L N(L,A)} n^{N(L,A)}}
\end{equation}

@@

Liouville theory, flat metric, asymptotic for field, Coulomb gas formalism.
\cite{nakayama2004liouville,ponsot2002boundary,teschner2001liouville,fateev2000boundary,teschner2000remarks,zamolodchikov1996conformal}

Liouville theory describes internal metric of surface. It is possible to obtain length of boundary
and area of surface with such a metric. To get the answer for flat loops we need to impose some
condition on metric in the bulk.

@@

In theory with the action
$$S=\frac{1}{4\pi} \int \sqrt{g}\left( \frac{1}{2} g^{\mu\nu} \partial_{\mu} \varphi \partial_{\nu}
  \varphi + Q R\varphi + 4\pi\mu e^{2b\varphi}\right) d^{2}x,$$ 
where $g$ is a background metric we need to assume some regularization procedure to define a
functional integral, because we want to integrate over measure that does not depend upon metric.
The idea is that the form of the action is the same but parameters are changed. 
So now $\mu$ is not a classical cosmological constant and $\mu,b,Q$ are connected through
consistency requirements for the theory. 
The consistency requirement is background-independence. The action should not change when we choose
another background metric and shift field $\varphi$ accordingly. 
Let $\tilde g = e^{\sigma} g$, redefine $\varphi$ as $\tilde \varphi=\varphi + \frac{Q}{2}\sigma$,
then all the terms with $\sigma$ are cancelled if we take into account contributions from matter
partition function. But it is possible only if 
$$Q=\frac{25-c_{M}}{6}$$. 
To make exponential term background-independent we need to have $Q=\frac{1}{b}+b$.

@@

@@

Here $g=\frac{1}{b^{2}}$. The result \eqref{eq:95} is supposed to coincide with the results of
\cite{fateev2000boundary} in the following sense:
\begin{equation}
  \label{eq:98}
  \partial_{\mu} W = \partial_{\mu_{B}} U
\end{equation}

>>>>

Some results exist in literature for particular case of self-avoiding walks ($\kappa=\frac{8}{3}$). We start
with the description of these results \ref{sec:self-avoid-polyg} and show that general problem
leads to Liouville field theory and obtain our result using Liouville
field theory fusion rules \ref{sec:liouville-theory}.

@@

Let us reformulate this problem a bit so that it could be generalized. Consider self-avoiding
polygons of length not greater than $n$ on a square lattice. This implies uniform measure on
polygons. The scaling limit of self-avoiding walk when lattice mesh size $a\to 0$ and length $l\to
\infty$ is conjectured to be described by Schramm-Loewner evolution with parameter
$\kappa=\frac{8}{3}$. Since we consider closed loops the limit is loops of conformal loop ensemble
with $\kappa=\frac{8}{3}$.
The natural question is how to generalize the definition of functions $G$ and $F$ to CLE with
arbitrary value of $\kappa$?

If we consider self-avoiding polygons of length up to $n$ the generating function $G_{n}(x)$ can be
introduced as non-local observable. Denote by $N_{n}$ total number of polygons of length up to $n$.
Since we have uniform measure, the probability for polygon to have length $l$ is
$\frac{p_{l}}{N_{n}}$. Then the generating function $G_{n}(x)$ is the expectation value of $x^{l}=e^{-\mu_{B}l}$
\begin{equation}
  \label{eq:102}
  G_{n}(x)=\sum_{l<n} p_{l} x^{l} = N_{n} \sum_{l<n} \frac{p_{l}}{N_{n}} x^{l} = N_{n}\mathbb{E}[x^{l}]
\end{equation}
Generating function for polygons of given area can be written in similar way as
\begin{equation}
  \label{eq:103}
   G_{n}(x,\mu)=\sum_{l<n} p_{l,A} x^{l} e^{-\mu A} = N_{n} \sum_{l<n} \frac{p_{l,A}}{N_{n}} x^{l} =
   N_{n}\mathbb{E}[x^{l} e^{-\mu A}]
\end{equation}
Note that the  observables $\mathbb{E}[e^{-\mu_{B}l}]$ ,$\mathbb{E}[e^{-\mu_{B}l -\mu A}]$ are
non-local. So while the generalization from self-avoiding walk to arbitrary SLE is natural it is not
clear how one can compute such observables.

Local observables w.r.t. SLE can be described by correlation functions of boundary conformal field
theory. So we need a theory which reproduces CFT in the limit $\mu,\mu_{B}\to 0$ and has non-local
observables such as area and length.

Liouville theory of two-dimensional gravity is a natural choice, but to see it we need to discuss
some geometry first.

Parameter $\kappa$ of SLE and CLE is connected with Hausdorff dimension of the curve.

@@

Let $N^r(L)$ be number of rooted loops with a given length. By $N^r(L,A)$ we denote the number of
loops of given length and area. $N^r(L)=\int_0^{\infty} N^r(L,A) dA$. Area weighted number of loops
is denoted by $G^r(L,\mu)$. (Here $\mu$ is bulk coupling constant ).
\begin{equation}
  \label{eq:2}
  G^r(L,\mu)=\int e^{\mu A} N^r(L,A) dA
\end{equation}

@@

%% 
%% \subsection{$O(n)$ model}
%% \label{sec:introduction-1}
%% 
%% Self-avoiding polygons can be thought as the limit $n\to 0$ of loops in $O(n)$ model.
%% 
%% Partition function for $O(n)$ model is
%% \begin{equation}
%%   \label{eq:9}
%%   Z_{O(n)} = \sum_{\mbox{loop conf}} x^{\mbox{total length}}
%%   n^{\mbox{number of loops}}
%% \end{equation}
%% Critical point is $x=\frac{1}{\sqrt{2+\sqrt{2-n}}}$, the continuum
%% limit in this point is described by minimal model $(p,p')$ with
%% $n=-2\cos\left(\frac{\pi p}{p'}\right)$. $c=1-6\frac{(p-p')^2}{pp'}$, $b=\sqrt{\frac{3}{2}}$.
%% 
%% Self-avoiding walk is obtained in the limit $n\to 0$. Then $c=0$ and
%% we have logarithmic conformal field theory. 
%% 
%% Logarithmic CFT $\sim$ Liouville theory ???
%% 
%% \section{Scaling of random loops}
%% \label{sec:scaling-random-loops}
%% 
%% Cardy \cite{cardy2001exact} argues that generating function $G^{(r)}$ for rooted self-avoiding loops is given by a derivative w.r.t. $x$
%% of expectation value of $e^{-gA}$ (eq.6):
%% \begin{equation}
%%   \label{eq:82}
%%   Z=\left<e^{-gA}\right>_{\mathrm{SAL}}=\left<e^{-\sqrt{g}\int J_{\lambda}A_{\lambda}d^{2}x}\right>_{\mathrm{SAL}.A}
%% \end{equation}
%% But reasoning is quite strange and unsound. Due to the need to use $g$ for metric in later sections
%% we will denote area coupling constant by $\mu$ and loop length coupling constant by $\mu_{B}=-\ln x$
%% (boundary). 
%% 
%% 
%% In this section we introduce scaling function for random loops. 
%% 
%% Consider rooted loops in $O(n)$ model.
%% 
%% The idea is to write $G$ as a correlation function in Liouville
%% theory.
%% 
%% 
%% Partition function is product over loops of different length
%% \begin{equation}
%%   \label{eq:10}
%%   Z_{O(n)} = \sum_{\mbox{loop conf}} \prod_L x^{L N(L)} n^{N(L)} = \sum_{\mbox{loop conf}} \prod_L\prod_A x^{L N(L,A)} n^{N(L,A)}
%% \end{equation}
%% 
%% 
%% We can rewrite $G^r(\mu,x)$ as the derivative of generating function
%% \begin{equation}
%%   \label{eq:7}
%%   G^r(\mu,x)=\mu\partial_x \log Z(\mu,\mu_B)
%% \end{equation}
%% 
%% Here
%% \begin{equation}
%%   \label{eq:8}
%%   Z(\mu,\mu_B) = \left< e^{-\mu A} \right> = \frac{\sum_{\mbox{loop conf}}
%%   \prod_L\prod_A x^{L N(L,A)} n^{N(L,A)} e^{-\mu A}}{\sum_{\mbox{loop conf}} \prod_L\prod_A x^{L N(L,A)} n^{N(L,A)}}
%% \end{equation}
%% 
%% 
KPZ-scaling. ``Wrong branch'' for $\mu_{B}$?

@@

%%  Liouville theory, flat metric, asymptotic for field, Coulomb gas formalism.
%%  \cite{nakayama2004liouville,ponsot2002boundary,teschner2001liouville,fateev2000boundary,teschner2000remarks,zamolodchikov1996conformal} 
%%  
%%  Liouville theory describes internal metric of surface. It is possible to obtain length of boundary
%%  and area of surface with such a metric. To get the answer for flat loops we need to impose some
%%  condition on metric in the bulk. 
%%

In last 30 years great progress was made in study of quantum Liouville theory made in papers
\cite{fateev2000boundary,zamolodchikov1996conformal,knizhnik1988fractal} and in study of matter on
random discrete surfaces
\cite{kazakov1986ising,duplantier1990geometrical,gaudin1989n,duplantier1988conformal}. These two
approaches to two-dimensional quantum gravity are in agreement
\cite{kostov2004boundary,kostov2003boundary}. In recent years strict mathematical definitions and proofs for some
field-theoretic results were obtained \cite{duplantier2011liouville,garban2012quantum}. All these
solid developments allow us to use Liouville theory as a theory of random surfaces and apply its
methods to the geometry of random loops. Our approach is similar to the application of Liouville
theory to harmonic measure problems \cite{duplantier2000conformally}.

In present text we do not try to be mathematically strict and do not use the definition of Liouville
measure from \cite{duplantier2011liouville} but we suppose that our result can be proven using
methods of that paper.

Before moving to quantum Liouville theory and scaling limit of generating functions of self-avoiding
polygons we need to discuss the geometry of Riemann surfaces and polygons on complex plane and its
connection with Liouville equation.

\subsection{Uniformization of Riemann surfaces and Liouville equation}
\label{sec:unif-riem-surf}

The problem of uniformization of Riemann surfaces was of central importance in mathematics at the
end of 19th -- beginning of 20th century. Great mathematicians such as Riemann, Klein and Poincare
proposed different approaches to its solution and the uniformization theorem was finally proven by
Koebe in 1907.

\begin{definition}
  A Riemann surface is a connected Hausdorff space $M$ (distinct points have disjoint neighbourhoods) together with a collection
  of charts ${U_{\alpha} , z_{\alpha} }$ with the following properties:
  \begin{itemize}
  \item The $U_{\alpha}$ form an open covering of $M$ .
  \item Each $z_{\alpha}$ is a homeomorphic mapping ($\exists$ continuous inverse) of $U_{\alpha}$ onto an open subset of the complex plane $\mathbb{C}$.
  \item If $U_{\alpha}\cap  U_{\beta} \neq \emptyset$, then $f_{\alpha\beta} = z_{\beta}\circ  z_{\alpha}^{-1}$ is complex analytic on $z_{\alpha} (U_{\alpha} \cap U_{\beta} )$.
  \end{itemize}  
\end{definition}

The statement of the theorem is 
\begin{theorem}
  Every simply connected Riemann surface is conformally equivalent to one of the three domains: the
  open unit disk, the complex plane, or the Riemann sphere. In particular it admits a Riemannian
  metric of constant curvature.
\end{theorem}

This classifies Riemannian surfaces as elliptic (positively curved -- rather, admitting a constant
positively curved metric), parabolic (flat), and hyperbolic (negatively curved) according to their
universal cover.

Proof is based upon existence and properties of Green function for the surface $M$.

 We can
 compose each $z_{\alpha}$ with a Mobius transformation of the disk onto itself so that we can
assume that for each coordinate disk $U_{\alpha}$ and each $p_{0}\in  U_{\alpha}$ there is a coordinate function
$z$ with $z (p_{0} ) = 0$.

     Suppose $M$ is a Riemann surface. Fix $p_{0}\in  M$ and let $z : U\to  \mathbb{D}$ be a coordinate
function such that $z(p_{0} ) = 0$. Let $\mathcal{F}_{p_{0}}$ be the collection of subharmonic
functions $v$ on $M \setminus p_{0}$ satisfying
\begin{itemize}
\item $v = 0$ on $M \setminus K$, for some compact $K\subset M$ with $K \neq M$, and
\item $\limsup_{p\to p_{0}} (v(p) + log |z(p)|) < \infty$
\end{itemize}

Set
\begin{equation}
  \label{eq:104}
  g_{M} (p, p_{0} ) = \sup\{v(p) : v\in \mathcal{F}_{p_{0}} \}
\end{equation}

Second condition does not depend on the choice of the coordinate function $z_{\alpha}$ , provided
$z_{\alpha} (p_{0} ) = 0$. The collection $\mathcal{F}_{p_{0}}$ is a Perron family, so by Harnack's
Theorem      $g_M (p, p_{0} )$ is either harmonic in  $M \setminus \{p_{0} \}$, or  is equal to $+\infty$  for
all $p\in M \setminus \{p_{0} \}$.
In the first case, $g_{M} (p, p_{0} )$ is called Green's function with pole (or logarithmic singularity)
at $p_{0}$ . In the second case we say that Green's function does not exist.

The Uniformization Theorem can be reformulated as follows (Koebe[1907]):

\begin{theorem}
  Suppose $M$ is a simply connected Riemann surface. If Green's function exists for $M$ , then there
  is a one-to-one analytic map of $M$ onto $\mathbb{D}$. If $M$ is compact, then there is a
  one-to-one analytic map of $M$ onto $\mathbb{C}^{*}$ . If $M$ is not compact and if Green's
  function does not exist for $M$ , then there is a one-to-one analytic map of $M$ onto
  $\mathbb{C}$.
\end{theorem}

Note that uniformization theorem does not give a construction of uniformization map.

More modern formulation of uniformization principle operates with quotient of universal cover of
Riemann surface by the action of some Fuchsian group. This group is monodromy group of Liouville
equation (?)

\subsection{Classical theory}
\label{sec:classical-theory}

@@

**** More on classical solutions here *****

\subsection{Functional integral formulation}
\label{sec:funct-integr-form}

To have quantum theory of gravity we need to define somehow functional integral. It can symbolically
written as
\begin{equation}
  \label{eq:101}
  \int [\mathcal{D}\varphi] e^{-S[\varphi]}
\end{equation}
Actual definition is tricky, because we need to quantize metric $e^{2b\varphi}g^{\mu\nu}$ and this
metric takes part not only in action for field $\varphi$, but also in action for matter field and
even functional integral measure \cite{david1988conformal,distler1989conformal,polyakov1987gauge}.

In theory with the action
$$S=\frac{1}{4\pi} \int \sqrt{g}\left( \frac{1}{2} g^{\mu\nu} \partial_{\mu} \varphi \partial_{\nu}
  \varphi + Q R\varphi + 4\pi\mu e^{2b\varphi}\right) d^{2}x,$$ 
where $g$ is a background metric we need to assume some regularization procedure to define a
functional integral, because we want to integrate over measure that does not depend upon metric.

The idea is that the form of the action is the same but parameters are changed. 
So now $\mu$ is not a classical cosmological constant and $\mu,b,Q$ are connected through
consistency requirements for the theory. 
The consistency requirement is background-independence. The action should not change when we choose
another background metric and shift field $\varphi$ accordingly. 
Let $\tilde g = e^{\sigma} g$, redefine $\varphi$ as $\tilde \varphi=\varphi + \frac{Q}{2}\sigma$,
then all the terms with $\sigma$ are cancelled if we take into account contributions from matter
partition function. But it is possible only if 
$$Q=\frac{25-c_{M}}{6}$$. 
To make exponential term background-independent we need to have $Q=\frac{1}{b}+b$.

@@

\subsubsection{KPZ-scaling}
\label{sec:kpz-scaling}

Knizhnik-Polyakov-Zamolodchikov (KPZ \cite{knizhnik1988fractal}) scaling relation gives behavior of
Liouville correlation functions w.r.t. to change in coupling constants $\mu, \mu_{B}$. This relation
is easy to obtain. Consider reparametrisation $\varphi\to\varphi-\frac{1}{2b}\ln\mu$. The action
$S[\mu,\mu_{B}]$ is then
\begin{equation}
  \label{eq:84}
  S[\mu,\mu_{B}]=-\frac{1}{4\pi}\int d^{2}x \frac{1}{2b}QR\ln\mu+S\left[1,\frac{\mu_{B}}{\sqrt{\mu}}\right]
\end{equation}
Since
\begin{equation}
  \label{eq:85}
  \int d^{2}x R =4\pi\chi,
\end{equation}
where $\chi=2-2\#handles-\#boundaries$ is Euler characteristic of the surface, we get
\begin{equation}
  \label{eq:99}
    S[\mu,\mu_{B}]=-\frac{1}{2b}Q\chi\ln\mu+S\left[1,\frac{\mu_{B}}{\sqrt{\mu}}\right]
\end{equation}
Substitute it into functional integral for correlation function of arbitrary bulk and boundary
vertex operators:
\begin{equation}
  \label{eq:100}
  \langle V_{\alpha_{1}}(z_{1})\dots V_{\alpha_{n}}(z_{n}) B_{\beta_{1}}(x_{1})\dots B_{\beta_{m}}(x_{m})\rangle=
\end{equation}

@@

Here $g=\frac{1}{b^{2}}$. The result \eqref{eq:95} is supposed to coincide with the results of
\cite{fateev2000boundary} \eqref{eq:21} in the following sense:
\begin{equation}
  \label{eq:98}
  \partial_{\mu} W = \partial_{\mu_{B}} U
\end{equation}
for $\alpha=b$ (?)

Substitute $  M^{2}=\mu\frac{\pi}{1}
\frac{\Gamma\left(\frac{1}{g}\right)}{\Gamma\left(1-\frac{1}{g}\right)}$ to \eqref{eq:95}:
\begin{equation}
  \label{eq:83}
  W(\tau) = -\frac{\left(\mu\frac{\pi}{1}
\frac{\Gamma\left(\frac{1}{g}\right)}{\Gamma\left(1-\frac{1}{g}\right)}\right)^{g/2}}{2\sin \pi g} \cosh g\tau,
\end{equation}

and compare with \eqref{eq:21} for $\alpha=b$:

\begin{equation}
  U(b)=\frac{2}{b} \left(\pi\mu\gamma( b^{2})\right)^{\frac{-b+1/b}{2b}} \Gamma(b^{2})
  \Gamma\left(1-\frac{1}{b^{2}}\right) \cosh (b-1/b)\pi s
\end{equation}
where $s$ is determined from the equation
\begin{equation}
  \cosh^{2} \pi b s=\frac{\mu_{B}^{2}}{\mu}\sin \pi b^{2}
\end{equation}
